Inefficient Concessions and Mediation

### 2023-08-17 draft_narrow.tex
change from Peace to Cooperation for stage-game outcome

In our model, concessions may act as signals or as an end in themselves -- or as both simultaneously. One contribution of this paper is that it allows concessions to have future material value, signaling value, and both. We will show that when concessions have both signaling and future material value, some of the material value of the concessions may be purposefully destroyed and thus inefficiencies can emerge. Another contribution is that the way we model concessions does not require countries to make an agreement that requires credible commitment, as is common in bargaining models. Instead, countries willingly provide a concession if they expect the signal it sends to provide a benefit in terms of future cooperation and public good provision.\footnote{When mediation is required to achieve peace in the model, it relies upon a manipulative mediator to enforce the required level of concessions.}

@@

Here $T>0$ represents the benefit from a country's negotiating partner playing Trust. $D>0$ represents the damages incurred when a country's negotiating partner plays Distrust, while $W>0$ is the additional benefit a country receives when it plays Distrust itself. We will refer to the stage-game Nash Equilibrium of (Distrust, Distrust) as ``War.'' We will refer to (Trust,Trust) as ``Peace.'' In line with the assumption that the countries play a Prisoner's Dilemma in the stage game, we assume $T>W-D$. Note that this ensures that the joint payoffs from ``Peace'' are the highest among all the stage-game action profiles.

@@

Parameters are common knowledge with the exception of $\delta_i$, which is country $i$'s private information. We take achieving peace, or more broadly, achieving the provision of public goods, as the normative goal with respect to advancing social welfare. Only high types are capable of making peace/providing public goods, and thus we will take the measure of social welfare to be the sum of the participating high types’ utilities.

\section{Analysis of Benchmark Cases}
\label{sec:bench}
When countries are faced with uncertainty about each other's motives, they can proceed in several ways, including giving concessions in order to signal their ability to make peace. We will examine three types of equilibria.\footnote{If the discount factor is high enough, there are many types of equilibria. Consider for instance oscillating every other period between (Trust,Trust) and (Distrust,Distrust) with a grim trigger ``War'' threat. In this paper, we are not interested in these equilibria because they are not as realistic for applications to conflict scenarios. These are also always payoff dominated by at least one of the other equilibria discussed and thus 	are not attractive.} The Appendix provides a thorough analysis of the various equilibria as well as most proofs.

@@

Without a third party, there is little hope of overcoming mistrust. These long lasting stalemates are the types of conflict which are most appropriate for mediator involvement. In Section~\ref{sec:med}, we will present a mediation mechanism that illustrates how mediation can help achieve peace in this context. In some sense, these are the cases that are most interesting because the international community might play a positive role.

\subsection{No-Concessions Separating Equilibrium}
\label{sec:ncse}
The other two equilibria on which we focus are separating equilibria in which low types and high types have different strategies. If $\delta_i$ is high enough for both countries, the ``Peace'' outcome can be sustained with no concessions and a grim trigger punishment threat. We call this a \emph{no-concessions separating equilibrium}. In this equilibrium, no concessions are given by either type in period 0. In period 1, high types play Trust while low types play Distrust. So long as there are two high types as revealed by their period 1 play, both countries play Trust in all following periods. Otherwise, both countries play Distrust in all following periods.

\begin{theorem}
\emph{No Concessions Separating Equilibrium}\\
	Assume $\delta_h \geq \frac{W}{(1-p)W + p(T+D)}=\de^{nc}$. A subgame perfect Nash equilibrium exists in which low types give no concessions in period 0 and always play Distrust and the high types give no concessions in period 0 and play Trust unless the other country defects to Distrust.
	\label{theorem:1}
\end{theorem}
\emph{Proof}: See Appendix.
\\
Here, as in each further iteration of the model, we define a high type as a country who has a discount factor at or above the threshold for cooperation. That is, high types prefer to play the ``Peace'' equilibrium, i.e. $\delta_h \geq \delta^{nc}$. Any country with $\delta_l < \delta^{nc}$ is a low type for the purposes of the \emph{no concessions separating equilibrium}. This means that two high types---and only two high types---can sustain the ``Peace'' outcome.\footnote{Note that, because we define high and low types relative to each type of equilibrium, a country with a given discount factor can be a high type in some equilibria and a low type in others.}

@@

\subsection{Concessions Separating Equilibrium}
\label{sec:cse}
The other separating equilibrium of interest involves distrustful countries using concessions as a way of building trust. We call this the \emph{concessions separating equilibrium}. In this equilibrium, the cost of giving a concession can be worth the investment if the other country also cooperates. It is also not as risky as the \emph{no-concession separating equilibrium}. In this equilibrium, concessions are given before more substantive actions and thus remove the possibility of playing one of the cheating corners of the Prisoner's Dilemma. Concessions are a costly signal that the other country is serious about cooperating and can be used to avoid continual distrust. That is, high types make use of concessions to identify themselves as trustworthy partners in hopes of establishing peace.

@@

Interestingly, the patience threshold for this equilibrium to exist, $\de^c = \frac{W}{T+D}$, is the same as the threshold for the complete information repeated game in which deviating from ``Trust'' is discouraged through grim trigger punishments.\footnote{When $\de = \frac{W}{T+D}$, there is also a semi-separating equilibrium where only some high types give a concession. Here, high types give a smaller expected concession and trade off a lower probability of achieving peace. We will thus focus on the pure concessions separating equilibrium in the analysis below because it provides the best chance for peace and is possible over a much larger portion of the parameter space.} Using concessions to separate the types therefore can be seen as making up for the imperfect information in this setting.

@@

Concessions separating equilibria are a way for countries to solve their disputes when the \emph{no concessions separating equilibrium} is unattractive or unavailable. By giving costly concessions, they signal their intentions in a way that a partner unwilling to make peace would not be willing to signal.

@@

Because $\al_i$ enters into the payoffs in a linear fashion, the choice that maximizes welfare in equilibrium is $\alpha=1$ for high types since a concession would only come from another high type in equilibrium, and thus only the ``Peace'' outcome will be played. Conversely, in equilibrium, low types only participate in the ``War'' equilibrium and thus maximize their payoffs by choosing $\alpha=0$.

@@

\begin{corollary}
	Assume that high types are not patient enough to separate without concessions but that they can separate using concessions when gifts have no material value. If countries use concessions with future material value instead of concessions with no future material value, there are parameters under which the future material value of gifts destroys countries' ability to separate through concessions, i.e. peace cannot be achieved when there are two high types.
	\label{corollary:matval}
\end{corollary}

@@

$\de_h$ is not involved in any of these calculations, so we can see that for any $\de_h \in [.5,.979)$, separating through concessions is possible when concessions have no material value but is not possible when they do have future material value. That is, future material value can destroy the opportunity for peace. \hfill $\blacksquare$

@@

Before turning to the question of how the possibility of destroying the value of concessions impacts the ability of the parties to achieve peace, we first formally establish the welfare implications.

\begin{lemma}
	If the \emph{concessions separating equilibrium with efficient concessions} exists, it is preferred to the \emph{concessions separating equilibrium with inefficient gifts} when both the probability of encountering a high type and the benefit from making peace are sufficiently large.
	\label{lemma:efficiency}
\end{lemma}
\emph{Proof:} See Appendix.

@@

This means that there are parameters under which ``Peace'' cannot be achieved in a \emph{concessions separating equilibrium with efficient gifts} but ``Peace'' can be achieved in a \emph{concessions separating equilibrium with inefficient gifts}. This is part of the substance of Theorem~\ref{theorem:4}.

@@

However, even if the negotiating parties are patient enough to separate when gifts are efficient, Lemma~\ref{lemma:efficiency} tells us that welfare will be higher with efficient gifts only when the probability of encountering a high type and the benefit from making peace are sufficiently large. Otherwise, welfare is higher under the \emph{concessions separating equilibrium with inefficient gifts} and this equilibrium is thus optimal. \hfill $\blacksquare$\\

Inefficient concessions are optimal when there's a good chance of facing a low type who will use the concession against you (i.e. $p$ is small) as well as when the benefits from ``Peace'' are small relative to the damages from ``War.'' This finding of inefficiency is for profoundly different reasons than the literature related to Camerer (1988) or Prendergast and Stole (2001).\footnote{Since our initial setup is essentially identical to Camerer's, inefficiency of  pre-play communication would also manifest in our model if a stage of pre-play communication were added. With a slight behavioral modification, the reasoning of Prendergast and Stole (2001) can also be had: if a country suffers regret after efficiently giving to a type different than their own, there is a premium for guessing correctly. In Prendergast and Stole the two types are equal except in labeling; in our model, high types and low types are used so there is an incentive for low types to try to pose as high types. However, putting this difference aside, the inefficiency result can be viewed as similar to that in Prendergast and Stole (2001). That is, if countries suffer a ``mismatching'' cost due to giving to the wrong type, they will be hesitant to do so and may offer inefficient concessions.} Instead of being in reaction to a behavioral regret for mismatching or serving as a way to discourage low types from entering the game altogether, here making concessions inefficient is a device to prevent low types from using those concessions against the giver of the concessions in the future.

@@

Ultimately, mediation is desirable if it can improve welfare or improve the chances for peace. Theorem~\ref{theorem:medwel} shows that a mediator who implements mechanism $M$ can do both.

\begin{theorem}
	When concessions have future material value, a mediator can use mechanism $M$ to
	\begin{enumerate}
		\item[(a)] improve welfare by increasing the efficiency of concessions when a \emph{concessions separating equilibrium with efficient gifts} is not possible; or
		\item[(b)] improve welfare despite reducing the efficiency of concessions when a \emph{concessions separating equilibrium with efficient gifts} is possible; or
		\item[(c)] achieve peace when it cannot be achieved either in a \emph{concessions separating equilibrium with efficient gifts} or a \emph{concessions separating equilibrium with inefficient gifts}.
	\end{enumerate}
	\label{theorem:medwel}
\end{theorem}
\emph{Proof:} See Appendix.

@@

Part (b) may be somewhat surprising. Here, a mediator would insert inefficiency when it is not necessary to make peace. This happens when the possibility of the low type using concessions against their negotiating partner and the magnitude of the damages that can be imposed are small enough that efficient gifts are possible but large enough that the full efficiency of the gifts has a large, negative impact on welfare. The mediator removes the possibility of the future damages, but has to require some inefficiency in order to incentivize low types to truthfully reveal.

Part (c) is relevant when the impacts of future material value are too strong, i.e when the probability of a low type is too high and/or the damages from investing in military power are too large. In this case, the parties cannot make peace on their own even with the possibility of signaling their type through concessions. By ensuring that only high types exchange gifts, the mediator removes the threat of damaging behavior by one's negotiating partner as well as the cost of giving the concession to a low type. Here, the help of the mediator allows types with lower patience levels to separate when those same types cannot separate in the bilateral game with concessions with future material value.

@@

 \noindent{\emph{Proof of Theorem \ref{theorem:1}}:\\
Playing Trust in the first period and thereafter as long as Trust has always been played will be incentive compatible for high types if the expected payoffs from this strategy are greater than the payoffs from playing Distrust in the first period and then every period thereafter by the one-shot deviation principle. If both countries play Trust in a period, a ``Peace'' equilibrium takes place and both countries continue to play Trust in every period with the grim trigger threat of the ``War'' equilibrium. Low types play Distrust in the first period, so a high type that observed Distrust being played would respond by playing Distrust in period 2 and in all future periods. A country plays Trust if and only if the following condition is satisfied:
\begin{equation*}
  p\left(\frac{T}{1- \delta}\right) +(1-p)\left({-D}+\frac{\delta(W-D)}{1- \delta} \right)\geq p \left(T + W + \frac{\de}{1-\de}(W-D)\right) +(1-p)\frac{W-D}{1- \delta}
\end{equation*}
Simplifying, we have
%\be
%  p\left(\frac{T}{1- \delta_h}\right) \geq p \left(T + W + \frac{\de_h}{1-\de_h}(W-D)\right) +(1-p)W
%\ee
\begin{equation*}
  pT \geq p \left((1-\de)(T + W) + \de (W-D)\right) + (1-\de)(1-p)W
\end{equation*}
\begin{equation*}
  pT \geq pT -p\de T +pW -p\de W + p \de W -p \de D  + W -\de W -pW + p\de W
\end{equation*}
\begin{equation*}
  p\de T +p \de D + \de W - p\de W  \geq W 
\end{equation*}
\be
	\delta \geq \frac{W}{(1-p)W + p(T+D)} = \de^{nc}
	\label{eq:ncse}
\ee

If the potential advantages of ``Peace'' are large enough to outweigh the risks of encountering a low type in expected utility terms, that is $\de_h \geq \de^{nc}$, a county has the incentive to play Trust in the first round rather than Distrust. Any player with $\de < \de^{nc}$ has the incentive---by this same inequality---to play Distrust in each period even if the other country plays Trust. Thus we call these countries low types for purposes of this equilibrium. \hfill $\blacksquare$

@@

High types have the incentive to play Trust only in the ``Peace'' equilibrium, which can only be sustained by a pair of high types. Both types benefit from the other country playing Trust so all countries have incentive to signal that they are high types if concessions are costless (cheap talk). Thus costless announcements cannot lead to truthful revelation. If concessions are to lead to a separating equilibrium, they must be costly. \hfill $\blacksquare$
\\
\\
\emph{Proof of Theorem \ref{theorem:2}:}\\
In the separating equilibrium constructed here, players will believe that any concession not equal to the equilibrium concession of the high type $g_h$ is a low type gift.  In this case, the binding incentive compatibility constraint is the low type IC constraint, which when simplified to allow $g_l=0$ following Lemma~\ref{lemma:3} is:\\  
\begin{equation*}
	pg_h +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g_h+pg_h 
\end{equation*}
This expression represents the low types' incentives for truth telling as opposed to posing as a high type and giving the corresponding equilibrium gift. Simplifying this inequality, we see that the high types need to give a gift 
\be
	g_h \geq p(X_{FT}^l-X_{FF}^l)
	\label{eq:cse}
\ee 
in order to separate from the low type.

@@

Here we specify that the mediator chooses $e=1$, which means that gift giving and receiving cancels out. The left hand side has zero gifts because the low-type player under consideration has truthfully revealed itself to be a low type and the mediator only specifies strictly positive gifts if there are two reports of `High.' Likewise, there are no gifts in the $(1-p)X_{FF}^l$ term on the right because the negotiating partner has made a report of `Low.' This leaves
		$$X_{FF}^l \geq X_{FT}^l.$$
		Expanding, we have
		\[
			\frac{W-D}{1-\de_l} \geq T + \left(\cdot \right) + W + \left(\cdot \right) + \frac{\de_l}{1-\de_l}\left(W+ \left(\cdot \right) -D - \left(\cdot \right) \right)
		\]
There are no gifts on the left hand side and so the payoffs from Table 1 apply. The right hand side has place holders $\left(\cdot\right)$ for the impact of the material value of gifts. On the right hand side, we have only the case where both parties declare themselves to be high types. Both parties are instructed to give efficient gifts. The negotiating partner really \emph{is} a high type and thus invests entirely in civil society because they expect to play the ``Peace'' equilibrium. The low-type player knows it will defect and be in a ``War'' equilibrium and so invests entirely in the military. 
%		\begin{multline}
%			W-D \geq 
%			\left(1-\de_l\right) \left[T + \al_2Tg^M + W + \left(1-\al_1\right)Wg_2 \right] \\
%			+ \de_l \left(W + \left(1-\al_1\right)Wg_2-D - \left(1-\al_2\right)Dg^M\right)
%			\label{eq:lowICex}
%		\end{multline}
		\begin{equation}
			W-D \geq 
			\left(1-\de_l\right) \left[T + Tg_h + W + Wg_h \right] + \de_l \left(W + Wg_h -D \right)
			\label{eq:med}
		\end{equation}
Canceling like terms and rearranging, we have
			\[
					0 \geq \left(1 - \de_l \right) T + \left(1 - \de_l \right) Tg_h + Wg_h + \left(1 - \de_l \right) D
			\]
			Since we assume throughout that gifts must be non-negative, each of $T, W$ and $D$ are positive and $0 \leq \de_l \leq 1$, this inequality can never hold.	\hfill $\blacksquare$\\
			\\
\emph{Proof of Theorem~\ref{theorem:med}:}\\
(a) The low type's individual rationality constraint is satisfied trivially: the low type gets the `War' payoff forever outside the mechanism; it also gets the `War' payoff forever \emph{inside} the mechanism while not giving or receiving any concession.

>>>>

In our model, concessions may act as signals or as an end in themselves -- or as both simultaneously. One contribution of this paper is that it allows concessions to have future material value, signaling value, and both. We will show that when concessions have both signaling and future material value, some of the material value of the concessions may be purposefully destroyed and thus inefficiencies can emerge. Another contribution is that the way we model concessions does not require countries to make an agreement that requires credible commitment, as is common in bargaining models. Instead, countries willingly provide a concession if they expect the signal it sends to provide a benefit in terms of future cooperation and public good provision.\footnote{When mediation is required to achieve cooperation in the model, it relies upon a manipulative mediator to enforce the required level of concessions.}

@@

Here $T>0$ represents the benefit from a country's negotiating partner playing Trust. $D>0$ represents the damages incurred when a country's negotiating partner plays Distrust, while $W>0$ is the additional benefit a country receives when it plays Distrust itself. We will refer to the stage-game Nash Equilibrium of (Distrust, Distrust) as ``War.'' We will refer to (Trust,Trust) as Cooperation. In line with the assumption that the countries play a Prisoner's Dilemma in the stage game, we assume $T>W-D$. Note that this ensures that the joint payoffs from Cooperation are the highest among all the stage-game action profiles.

@@

Parameters are common knowledge with the exception of $\delta_i$, which is country $i$'s private information. We take achieving peace, or more broadly, achieving cooperation with the provision of public goods, as the normative goal with respect to advancing social welfare. Only high types are capable of making peace/providing public goods, and thus we will take the measure of social welfare to be the sum of the participating high types’ utilities.

\section{Analysis of Benchmark Cases}
\label{sec:bench}
When countries are faced with uncertainty about each other's motives, they can proceed in several ways, including giving concessions in order to signal their ability to Cooperate. We will examine three types of equilibria.\footnote{If the discount factor is high enough, there are many types of equilibria. Consider for instance oscillating every other period between (Trust,Trust) and (Distrust,Distrust) with a grim trigger ``War'' threat. In this paper, we are not interested in these equilibria because they are not as realistic for applications to conflict scenarios. These are also always payoff dominated by at least one of the other equilibria discussed and thus 	are not attractive.} The Appendix provides a thorough analysis of the various equilibria as well as most proofs.

@@

Without a third party, there is little hope of overcoming mistrust. These long lasting stalemates are the types of conflict which are most appropriate for mediator involvement. In Section~\ref{sec:med}, we will present a mediation mechanism that illustrates how mediation can help achieve cooperation in this context. In some sense, these are the cases that are most interesting because the international community might play a positive role.

\subsection{No-Concessions Separating Equilibrium}
\label{sec:ncse}
The other two equilibria on which we focus are separating equilibria in which low types and high types have different strategies. If $\delta_i$ is high enough for both countries, the Cooperation outcome can be sustained with no concessions and a grim trigger punishment threat. We call this a \emph{no-concessions separating equilibrium}. In this equilibrium, no concessions are given by either type in period 0. In period 1, high types play Trust while low types play Distrust. So long as there are two high types as revealed by their period 1 play, both countries play Trust in all following periods. Otherwise, both countries play Distrust in all following periods.

\begin{theorem}
\emph{No Concessions Separating Equilibrium}\\
	Assume $\delta_h \geq \frac{W}{(1-p)W + p(T+D)}=\de^{nc}$. A subgame perfect Nash equilibrium exists in which low types give no concessions in period 0 and always play Distrust and the high types give no concessions in period 0 and play Trust unless the other country defects to Distrust.
	\label{theorem:1}
\end{theorem}
\emph{Proof}: See Appendix.
\\
Here, as in each further iteration of the model, we define a high type as a country who has a discount factor at or above the threshold for cooperation. That is, high types prefer to play the Cooperation equilibrium, i.e. $\delta_h \geq \delta^{nc}$. Any country with $\delta_l < \delta^{nc}$ is a low type for the purposes of the \emph{no concessions separating equilibrium}. This means that two high types---and only two high types---can sustain the Cooperation outcome.\footnote{Note that, because we define high and low types relative to each type of equilibrium, a country with a given discount factor can be a high type in some equilibria and a low type in others.}

@@

\subsection{Concessions Separating Equilibrium}
\label{sec:cse}
The other separating equilibrium of interest involves distrustful countries using concessions as a way of building trust. We call this the \emph{concessions separating equilibrium}. In this equilibrium, the cost of giving a concession can be worth the investment if the other country also cooperates. It is also not as risky as the \emph{no-concession separating equilibrium}. In this equilibrium, concessions are given before more substantive actions and thus remove the possibility of playing one of the cheating corners of the Prisoner's Dilemma. Concessions are a costly signal that the other country is serious about cooperating and can be used to avoid continual distrust. That is, high types make use of concessions to identify themselves as trustworthy partners in hopes of establishing cooperation.

@@

Interestingly, the patience threshold for this equilibrium to exist, $\de^c = \frac{W}{T+D}$, is the same as the threshold for the complete information repeated game in which deviating from ``Trust'' is discouraged through grim trigger punishments.\footnote{When $\de = \frac{W}{T+D}$, there is also a semi-separating equilibrium where only some high types give a concession. Here, high types give a smaller expected concession and trade off a lower probability of achieving cooperation. We will thus focus on the pure concessions separating equilibrium in the analysis below because it provides the best chance for Cooperation and is possible over a much larger portion of the parameter space.} Using concessions to separate the types therefore can be seen as making up for the imperfect information in this setting.

@@

Concessions separating equilibria are a way for countries to solve their disputes when the \emph{no concessions separating equilibrium} is unattractive or unavailable. By giving costly concessions, they signal their intentions in a way that a partner unwilling to cooperate would not be willing to signal.

@@

Because $\al_i$ enters into the payoffs in a linear fashion, the choice that maximizes welfare in equilibrium is $\alpha=1$ for high types since a concession would only come from another high type in equilibrium, and thus only the Cooperation outcome will be played. Conversely, in equilibrium, low types only participate in the ``War'' equilibrium and thus maximize their payoffs by choosing $\alpha=0$.

@@

\begin{corollary}
	Assume that high types are not patient enough to separate without concessions but that they can separate using concessions when gifts have no material value. If countries use concessions with future material value instead of concessions with no future material value, there are parameters under which the future material value of gifts destroys countries' ability to separate through concessions, i.e. Cooperation cannot be achieved when there are two high types.
	\label{corollary:matval}
\end{corollary}

@@

$\de_h$ is not involved in any of these calculations, so we can see that for any $\de_h \in [.5,.979)$, separating through concessions is possible when concessions have no material value but is not possible when they do have future material value. That is, future material value can destroy the opportunity for cooperation. \hfill $\blacksquare$

@@

Before turning to the question of how the possibility of destroying the value of concessions impacts the ability of the parties to cooperate, we first formally establish the welfare implications.

\begin{lemma}
	If the \emph{concessions separating equilibrium with efficient concessions} exists, it is preferred to the \emph{concessions separating equilibrium with inefficient gifts} when both the probability of encountering a high type and the benefit from Cooperation are sufficiently large.
	\label{lemma:efficiency}
\end{lemma}
\emph{Proof:} See Appendix.

@@

This means that there are parameters under which Cooperation cannot be achieved in a \emph{concessions separating equilibrium with efficient gifts} but Cooperation can be achieved in a \emph{concessions separating equilibrium with inefficient gifts}. This is part of the substance of Theorem~\ref{theorem:4}.

@@

However, even if the negotiating parties are patient enough to separate when gifts are efficient, Lemma~\ref{lemma:efficiency} tells us that welfare will be higher with efficient gifts only when the probability of encountering a high type and the benefit from Cooperation are sufficiently large. Otherwise, welfare is higher under the \emph{concessions separating equilibrium with inefficient gifts} and this equilibrium is thus optimal. \hfill $\blacksquare$\\

Inefficient concessions are optimal when there's a good chance of facing a low type who will use the concession against you (i.e. $p$ is small) as well as when the benefits from Cooperation are small relative to the damages from ``War.'' This finding of inefficiency is for profoundly different reasons than the literature related to Camerer (1988) or Prendergast and Stole (2001).\footnote{Since our initial setup is essentially identical to Camerer's, inefficiency of  pre-play communication would also manifest in our model if a stage of pre-play communication were added. With a slight behavioral modification, the reasoning of Prendergast and Stole (2001) can also be had: if a country suffers regret after efficiently giving to a type different than their own, there is a premium for guessing correctly. In Prendergast and Stole the two types are equal except in labeling; in our model, high types and low types are used so there is an incentive for low types to try to pose as high types. However, putting this difference aside, the inefficiency result can be viewed as similar to that in Prendergast and Stole (2001). That is, if countries suffer a ``mismatching'' cost due to giving to the wrong type, they will be hesitant to do so and may offer inefficient concessions.} Instead of being in reaction to a behavioral regret for mismatching or serving as a way to discourage low types from entering the game altogether, here making concessions inefficient is a device to prevent low types from using those concessions against the giver of the concessions in the future.

@@

Ultimately, mediation is desirable if it can improve welfare or improve the chances for cooperation. Theorem~\ref{theorem:medwel} shows that a mediator who implements mechanism $M$ can do both.

\begin{theorem}
	When concessions have future material value, a mediator can use mechanism $M$ to
	\begin{enumerate}
		\item[(a)] improve welfare by increasing the efficiency of concessions when a \emph{concessions separating equilibrium with efficient gifts} is not possible; or
		\item[(b)] improve welfare despite reducing the efficiency of concessions when a \emph{concessions separating equilibrium with efficient gifts} is possible; or
		\item[(c)] achieve cooperation when it cannot be achieved either in a \emph{concessions separating equilibrium with efficient gifts} or a \emph{concessions separating equilibrium with inefficient gifts}.
	\end{enumerate}
	\label{theorem:medwel}
\end{theorem}
\emph{Proof:} See Appendix.

@@

Part (b) may be somewhat surprising. Here, a mediator would insert inefficiency when it is not necessary to achieve a cooperative outcome. This happens when the possibility of the low type using concessions against their negotiating partner and the magnitude of the damages that can be imposed are small enough that efficient gifts are possible but large enough that the full efficiency of the gifts has a large, negative impact on welfare. The mediator removes the possibility of the future damages, but has to require some inefficiency in order to incentivize low types to truthfully reveal.

Part (c) is relevant when the impacts of future material value are too strong, i.e when the probability of a low type is too high and/or the damages from investing in military power are too large. In this case, the parties cannot cooperate on their own even with the possibility of signaling their type through concessions. By ensuring that only high types exchange gifts, the mediator removes the threat of damaging behavior by one's negotiating partner as well as the cost of giving the concession to a low type. Here, the help of the mediator allows types with lower patience levels to separate when those same types cannot separate in the bilateral game with concessions with future material value.

@@

 \noindent{\emph{Proof of Theorem \ref{theorem:1}}:\\
Playing Trust in the first period and thereafter as long as Trust has always been played will be incentive compatible for high types if the expected payoffs from this strategy are greater than the payoffs from playing Distrust in the first period and then every period thereafter by the one-shot deviation principle. If both countries play Trust in a period, a Cooperation equilibrium takes place and both countries continue to play Trust in every period with the grim trigger threat of the ``War'' equilibrium. Low types play Distrust in the first period, so a high type that observed Distrust being played would respond by playing Distrust in period 2 and in all future periods. A country plays Trust if and only if the following condition is satisfied:
\begin{equation*}
  p\left(\frac{T}{1- \delta}\right) +(1-p)\left({-D}+\frac{\delta(W-D)}{1- \delta} \right)\geq p \left(T + W + \frac{\de}{1-\de}(W-D)\right) +(1-p)\frac{W-D}{1- \delta}
\end{equation*}
Simplifying, we have
%\be
%  p\left(\frac{T}{1- \delta_h}\right) \geq p \left(T + W + \frac{\de_h}{1-\de_h}(W-D)\right) +(1-p)W
%\ee
\begin{equation*}
  pT \geq p \left((1-\de)(T + W) + \de (W-D)\right) + (1-\de)(1-p)W
\end{equation*}
\begin{equation*}
  pT \geq pT -p\de T +pW -p\de W + p \de W -p \de D  + W -\de W -pW + p\de W
\end{equation*}
\begin{equation*}
  p\de T +p \de D + \de W - p\de W  \geq W 
\end{equation*}
\be
	\delta \geq \frac{W}{(1-p)W + p(T+D)} = \de^{nc}
	\label{eq:ncse}
\ee

If the potential advantages of Cooperation are large enough to outweigh the risks of encountering a low type in expected utility terms, that is $\de_h \geq \de^{nc}$, a county has the incentive to play Trust in the first round rather than Distrust. Any player with $\de < \de^{nc}$ has the incentive---by this same inequality---to play Distrust in each period even if the other country plays Trust. Thus we call these countries low types for purposes of this equilibrium. \hfill $\blacksquare$

@@

High types have the incentive to play Trust only in the Cooperation equilibrium, which can only be sustained by a pair of high types. Both types benefit from the other country playing Trust so all countries have incentive to signal that they are high types if concessions are costless (cheap talk). Thus costless announcements cannot lead to truthful revelation. If concessions are to lead to a separating equilibrium, they must be costly. \hfill $\blacksquare$
\\
\\
\emph{Proof of Theorem \ref{theorem:2}:}\\
In the separating equilibrium constructed here, players will believe that any concession not equal to the equilibrium concession of the high type $g_h$ is a low type gift.  In this case, the binding incentive compatibility constraint is the low type IC constraint, which when simplified to allow $g_l=0$ following Lemma~\ref{lemma:3} is:\\  
\begin{equation*}
	pg_h +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g_h+pg_h 
\end{equation*}
This expression represents the low types' incentives for truth telling as opposed to posing as a high type and giving the corresponding equilibrium gift. Simplifying this inequality, we see that the high types need to give a gift 
\be
	g_h \geq p(X_{FT}^l-X_{FF}^l)
	\label{eq:cse}
\ee 
in order to separate from the low type.

@@

Here we specify that the mediator chooses $e=1$, which means that gift giving and receiving cancels out. The left hand side has zero gifts because the low-type player under consideration has truthfully revealed itself to be a low type and the mediator only specifies strictly positive gifts if there are two reports of `High.' Likewise, there are no gifts in the $(1-p)X_{FF}^l$ term on the right because the negotiating partner has made a report of `Low.' This leaves
		$$X_{FF}^l \geq X_{FT}^l.$$
		Expanding, we have
		\[
			\frac{W-D}{1-\de_l} \geq T + \left(\cdot \right) + W + \left(\cdot \right) + \frac{\de_l}{1-\de_l}\left(W+ \left(\cdot \right) -D - \left(\cdot \right) \right)
		\]
There are no gifts on the left hand side and so the payoffs from Table 1 apply. The right hand side has place holders $\left(\cdot\right)$ for the impact of the material value of gifts. On the right hand side, we have only the case where both parties declare themselves to be high types. Both parties are instructed to give efficient gifts. The negotiating partner really \emph{is} a high type and thus invests entirely in civil society because they expect to play the Cooperation equilibrium. The low-type player knows it will defect and be in a ``War'' equilibrium and so invests entirely in the military. 
%		\begin{multline}
%			W-D \geq 
%			\left(1-\de_l\right) \left[T + \al_2Tg^M + W + \left(1-\al_1\right)Wg_2 \right] \\
%			+ \de_l \left(W + \left(1-\al_1\right)Wg_2-D - \left(1-\al_2\right)Dg^M\right)
%			\label{eq:lowICex}
%		\end{multline}
		\begin{equation}
			W-D \geq 
			\left(1-\de_l\right) \left[T + Tg_h + W + Wg_h \right] + \de_l \left(W + Wg_h -D \right)
			\label{eq:med}
		\end{equation}
Canceling like terms and rearranging, we have
			\[
					0 \geq \left(1 - \de_l \right) T + \left(1 - \de_l \right) Tg_h + Wg_h + \left(1 - \de_l \right) D
			\]
			Since we assume throughout that gifts must be non-negative, each of $T, W$ and $D$ are positive and $0 \leq \de_l \leq 1$, this inequality can never hold.	\hfill $\blacksquare$\\
			\\
\emph{Proof of Theorem~\ref{theorem:med}:}\\
(a) The low type's individual rationality constraint is satisfied trivially: the low type gets the `War' payoff forever outside the mechanism; it also gets the `War' payoff forever \emph{inside} the mechanism while not giving or receiving any concession.
